---
title: "Lab 21: Sixteen Personality Factors I"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Sixteen Personality Factors: Pick Your Trait

Over the next two classes we are going to be looking at survey response
data from the Sixteen Personality Factor Questionnaire in order to practice
our skills at statistical inference:

```{r}
pf <- read_csv("https://statsmaths.github.io/stat_data/cattell_16pf.csv")
```

The dataset use the following fields:

- age: respondent's age in years
- gender: respondent's self-selected gender
- country: two letter IATA code for the respondent's IP
- elapsed: time taken to complete quiz in seconds
- warmth: personality score from 1-20
- reasoning: personality score from 1-20
- emotional_stability: personality score from 1-20
- dominance: personality score from 1-20
- liveliness: personality score from 1-20
- rule_consciousness: personality score from 1-20
- social_boldness: personality score from 1-20
- sensitivity: personality score from 1-20
- vigilance: personality score from 1-20
- abstractedness: personality score from 1-20
- privateness: personality score from 1-20
- apprehension: personality score from 1-20
- openness_to_change: personality score from 1-20
- self_reliance: personality score from 1-20
- perfectionism: personality score from 1-20
- tension: personality score from 1-20
- baseline: average score across all 16 personality traits

*To start with, select a particular trait that you will use for the first
bank of questions. I suggest picking something that popped out when you took
the test. You can pick any of the 16 other than sensitivity.*

Produce a bar plot of the personality scores for your trait.
```{r}
ggplot(pf, aes(emotional_stability)) +
  geom_bar()
```

Describe the distribution. Why does a bar plot work here even though the
variable is numeric?

**Answer**: The graph shows that there is not an even distribution, most of the people have an emotional stability score between 10-15. It workds becuase the numerical data is also categorical, it can only be between 1-20. 

Produce a confidence interval for the mean of your trait.

```{r}
model <- lm_basic(emotional_stability ~ 1, data = pf)
reg_table(model, level= 0.95)
```

Now, produce a dataset that consists only of responses from the country of
Hungary. The 2-letter country code for Hungary is "HU". Produce a confidence
interval for your trait on the Hungarian dataset.

```{r}
hungary <- filter(pf, country == "HU")
hungary

hungarymodel <- lm_basic(emotional_stability ~ 1, data = hungary)
reg_table(hungarymodel, level=0.95)
```

What do you notice about this confidence interval compared to the original
one? Can you explain why this is the case?

**Answer**: The confidence interval for the Hungary data is much bigger than the confidence interval for the original data. I think this is the case because the original data encompasses more observations therefore one particular score or scores will not be seen as "outliers".

Find a 95% confidence interval for the difference between the average
male and female value for your personality trait on the Hungarian data.

```{r}
hungarymodel <- lm_basic(emotional_stability ~ 1 + gender, data = hungary)
reg_table(hungarymodel, level=0.95)
```


**Answer**: The average female emtional stability score is 12.28 and the average male emotional stability score is 12.17. 

Is there a statistically significant difference between men and women's
average score? If so, what direction is this difference? Does this challenge
or confirm traditional gender stereotypes (note: not all personality traits
have one)?

**Answer**: No, there is not a statistically significant difference. 

Construct a new dataset that only has ages from 30-49 (from the original `pf`,
not just the Hungarian subset). Hint: you can use the filter function twice.

```{r}
middleage <- filter(pf, age <= 49)
middleage <- filter(middleage, age >= 30)
```

On the dataset of people ages 30 to 49, create a variable called `fourties` if
age is greater than or equal to 40.

```{r}
middleage$fourties <- (middleage$age >= 40)
```


Fit a regression on with your personality trait as a response to test for the
difference between its mean for people in their 30s versus people in their
40s. Compute a confidence interval for this difference.

```{r}
middleagemodel <- lm_basic(emotional_stability ~ 1 + fourties, data = middleage)
reg_table(middleagemodel, level = 0.95)
```

Is there a statistically significant difference between 30s and 40s average
scores? If so, what direction is this difference? Does this challenge or
confirm traditional age stereotypes (note: not all personality traits
have one)?

**Answer**: No, there is not.

Draw a bar plot of the variable `country` over the whole dataset. Take
note of the number of countries with a very small number of responses.

```{r}
ggplot(pf, aes(country)) +
  geom_bar()
```

Fitting a model with all of these countries is possible but not very
useful. When we only have a few responses from some places it makes those
few regions with a lot of data harding to identify and analyze. Fortunately
there is a solution in the function `fct_lump`.

Fit a regression model on the entire dataset by the country variable lumped
into 5 categories.

```{r}
countrymodel <- lm_basic(emotional_stability ~ fct_lump(country, n = 5), data = pf)
reg_table(countrymodel, level = 0.95)


```

Using the previous model, which country (not including "Other") has the
highest score for your trait? Which has the lowest?

**Answer**: The country that has the highest score is IN, the country that has the lowest score is the US.

## Fitting Sensitivity Scores

Fit a model that predicts sensitivity as a function of the baseline
score.

```{r}
sensitivitymodel <- lm_basic(sensitivity ~ 1 + baseline, data = pf)
reg_table(sensitivitymodel, level = 0.95)
```

Is the slope statistically significantly different from 1? Why is this an
interesting question in the context of the data?

**Answer**: No it is not. Maybe it shows that the sensitivity score was not significant in this data set since the baseline score and .

Fit a regression model that uses both gender and the baseline score to
predict your trait.

```{r}
triplemodel <- lm_basic(emotional_stability ~ 1 + baseline + gender, data = pf)
reg_table(triplemodel, level= 0.95)
```

Describe the slope for the gender term in words:

**Answer**: The slope shows the average emotional stability score by gender. The slope for the emotional stability of females is higher than the slope for males. 

Add predictions from the previous model to `pf`.

```{r}
pf <- add_prediction(pf, triplemodel)
pf
```

Plot the baseline score as a function of `model_pred`, coloring the points
base on the gender variable.

```{r}
ggplot(pf, aes(emotional_stability, baseline)) +
  geom_point() +
  geom_point(aes( y = triplemodel_pred, color = gender)) +
  geom_line( aes(y = triplemodel_pred, color = gender))
```

What do the predicted values look like?

**Answer**: The predicted values all follow the same trend but are scattered within a constrained area.


