---
title: "Lab 18: Inference for the Mean"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Teacher Salary

We will start today by looking at a small dataset containing teacher
salaries from 2009-2010 for 71 randomly choosen teachers employed by the
St. Louis Public School in in Michigan.

```{r}
teachers <- read_csv("https://statsmaths.github.io/stat_data/teachers_pay.csv")
```

The available variables are

- base: Base annual salary, in dollars
- degree: Highest educational degree attained: BA (Bachelor's) or MA (Master's)
- years: Number of years employeed

Using the mean function, what is the average base pay of all teachers
in the dataset?

```{r}
mean(teachers$base)
```

Fit a model for the mean of the base pay variable using lm_basic. Save
the model as an object called "model":

```{r}
model <- lm_basic(base ~ 1, data = teachers)
model
```

Using a call to `reg_table`, find the mean implied by the model:

```{r}
reg_table(model)
```

Does the mean agree with your answer to question 2?

**Answer**: Yes, it just rounded it up to the nearest dollar.

Add a 95% confidence interval to the regression table.

```{r}
reg_table(model, level = 0.95)
```

What is the range of mean salaries implied by the confidence interval?

**Answer**: The range is $54,785 to $59,091.

Draw a histogram of the base salary values for the entire dataset.

```{r}
ggplot(teachers, aes(base)) +
  geom_histogram()
```

Do most of the salary values fall within the range given in question 5? Why
or why not?

**Answer**: No, because the range in question five was to narrow down where approximately the mean was going to fall.  

Use the filter command to construct a new dataset called `masters` consisting
of just those teachers with a masters degree.

```{r}
masters <- filter(teachers, degree == "MA")
masters
```

Compute a 95% confidence interval for the mean pay of teachers with a master's
degree. Does this range intersect the one you had in question 5?

```{r}
mastersmodel <- lm_basic(base ~ 1, data = masters)
mastersmodel

reg_table(mastersmodel, level = 0.95)
```

The range for the Master's degree lies about %98 within the range form question five. The other 2% lies outside of the higher end of the range.

## Murder Data

Now load the following dataset containing all murders that have occurred in
London from 1 January 2006 to 7 September 7 2011.

```{r}
london <- read_csv("https://statsmaths.github.io/stat_data/london_murders.csv")
```

The available variables are:

- age: age of the victim in years
- year: year of the murder
- borough: the London borough in which the murder took place

Find an 80% confidence interval for the average age of the victim of a
murder in London.

```{r}
murderage <- lm_basic(age ~ 1, data = london)
murderage

reg_table(murderage, level = 0.80)
```

Make sure you actually extract the answer here:

**Answer**: The confidence interval for the average age of a victim is between 33 years and 34 years. 

Describe in words how the proceeding confidence interval should be
interpreted.

**Answer**: The confidence interval above shows the range in which the median should fall and that is how it should be read. It is the range for average not for the entire dataset.








